---
title: "01 Packages and functions"
output: html_notebook
---

# Packages

*All* the packages needed for this project should be loaded here so they can be
easily reloaded after a restart.

```{r}
# circular statistics
library("circular")
# spatial processing
library("sf")
# tidyverse loads dplyr, forcats, ggplot2, purrr, readr, stringr, tibble, tidyr
library("tidyverse")
# repelling labels in ggplot2
library("ggrepel")
# date processing
library("lubridate")
# tidy model results
library("broom")
```


# Parameters

```{r}
cities <- tribble(
  ~name, ~fips, ~prefix,
  "Chicago",        "17", "chi",
  "Detroit",        "26", "dtt",
  "Fort Worth",     "48", "ftw",
  "Kansas City",    "29", "kcm",
  "Los Angeles",    "06", "lax",
  "Louisville",     "21", "lou",
  "New York",       "36", "nyc",
  "San Francisco",  "06", "sfo",
  "Tucson",         "04", "tus",
  "Virginia Beach", "51", "vib"
)
```



# Functions

## Format a p value

```{r}
format_p <- function (p) {
  ifelse(p < 0.001, "<0.001", sprintf("%3.3f", p))
}
```


## Sample vector only if long enough

`sample_n()` throws an error if the table to be sampled from has fewer rows than
the value specified for n. This function deals with that by only sampling 
tables that are longer than the required sample size, otherwise returning the
table unchanged. The same behaviour is provided for vectors, which sample_n
does not handle.

```{r}
sample_if <- function (x, n) {
  
  if (is.atomic(x)) {
    
    # vector
    if (length(x) > n) {
      sample(x, n)
    } else {
      x
    }
    
  } else {
    
    # table
    if (nrow(x) > n) {
      sample_n(x, n)
    } else {
      x
    }
    
  }
}
```


## Convert vector of time angles to circular object

This function takes a vector of angles, i.e. numeric values between 0 and 360
degrees, and converts them into a circular object with appropriate attributes.

```{r}
circular_clock <- function (x) {
  if (!is.numeric(x)) {
    stop("x must be numeric, not of class ", class(x))
  }
  circular(x, units = "degrees", zero = pi / 2, rotation = "clock")
}
```


## Convert negative angles to their positive equivalents

This function takes a numeric vector of angles and replaces those that are
negative with the appropriate positive value. For example, -90ยบ is the same
angle as +270ยบ.

```{r}
abs_circular <- function (x) {
  if (!is.numeric(x)) {
    stop("x must be numeric, not of class ", class(x))
  }
  ifelse(x < 0, x + 360, x)
}
```


## Calculate circular concentration

```{r}
circular_concentration <- function (x) {
  quantile.circular(x, probs = c(0.25, 0.75)) %>% 
    abs_circular() %>% 
    { ifelse(.[2] > .[1], .[2] - .[1], .[1] - .[2]) } %>% 
    { 1 - ((ifelse(. > 180, 360 - ., .) / 360) / 0.5) }
    # { 1 - ((ifelse(.[2] - .[1] > 180, 360 - (.[2] - .[1]), .) / 360) / 0.5) }
}
```


## Calculate seasonal range

This function accepts a vector of date objects and calculates the 10% trimmed
range of the daily/weekly/monthly count of offenses, or the range excluding the
minimum and maximum values if there are fewer than 10 periods.

```{r}
seasonal_range <- function (
  x, # a vector of date objects
  period, # the period over which offenses should be counted
  ratio = TRUE, # whether to return the result divided by the median period count
  edges = "exclude" # how edges should be dealt with if period = "weeks"
) {
  
  # check arguments
  if (!is.POSIXt(x) & !is.Date(x)) {
    stop("x must be a Date, POSIXct or POSIXlt, not of class ", class(x))
  }
  if (!edges %in% c("exclude", "combine")) {
    stop("edges must be one of 'exclude' (the default) or 'combine'")
  }
  
  # if period is a single letter, convert it into a format code
  if (length(period) == 1 & period %in% c(letters, LETTERS)) {
    period <- paste0("%", period)
  }
  
  # calculate offense counts
  y <- tibble(date = x) %>% 
    mutate(date_part = format(date, format = period)) %>% 
    group_by(date_part) %>% 
    summarise(n = n())
  
  # deal with edges in weeks
  if (period %in% c("%U", "%V", "%W")) {
    if (edges == "exclude") {
      y <- filter(y, !date_part %in% c("00", "53"))
    } else if (edges == "combine") {
      y <- mutate(y, date_part = case_when(
        date_part == "53" ~ "01",
        date_part == "00" ~ "52",
        TRUE ~ date_part
      ))
    }
  }

  # calculate min and max
  if (nrow(y) == 1) {
    stop("all the values of x are contained in a single period so a range ",
         "cannot be calculated")
  } else if (nrow(y) < 10) {
    z <- y %>% arrange(n) %>% slice(2:(nrow(.) - 1)) %>% { range(.$n) }
  } else {
    z <- quantile(y$n, probs = c(0.1, 0.9))
  }
  
  # calculate range
  if (ratio == TRUE) {
    as.numeric(z[2] - z[1]) / median(y$n)
  } else {
    as.numeric(z[2] - z[1])
  }
  
}
```


## Angle to date

```{r}
angle_to_day <- function (x, format = "%d %b") {
  floor((x / 360) * 365.25) %>% 
    as.Date(origin = "2010-01-01") %>% 
    as.character(format = format)
}
```


## Angle to number of days

```{r}
angle_to_days <- function (x) {
  round((x / 360) * 365.25)
}
```


