---
title: "02 Data preparation"
output: html_notebook
---


# Load CODE crime data

```{r}
crimes <- dir("../original_data", pattern = "_core_", full.names = TRUE) %>% 
  map(read_csv, col_types = cols(
    .default = col_character(),
    uid = col_integer(),
    date_single = col_datetime(format = ""),
    date_start = col_datetime(format = ""),
    date_end = col_datetime(format = ""),
    longitude = col_double(),
    latitude = col_double()
  )) %>% 
  bind_rows() %>% 
  # Virginia Beach data is only available for four years, so is excluded
  filter(city_name != "Virginia Beach") %>% 
  # not all fields are needed, so some a excluded to save memory
  select(uid, city_name, offense_code, offense_type, offense_group, 
         offense_against, date_single, date_start, date_end, longitude, latitude, 
         location_type, location_category, fips_state, fips_county, tract, 
         block_group, block)

crimes %>% 
  mutate(year = year(date_single)) %>% 
  group_by(city_name, year) %>% 
  summarise(n = n()) %>% 
  spread(year, n)
```


# Create circular representations of dates

Circular statistics require dates to be represented as angles showing the 
position of that date within a larger unit. For example, an angle could show
how far through the year, week, day etc a date/time combination falls. Since
we're interested in seasonal variations, we will represent dates as an angle
relative to a year.

```{r}
crimes <- crimes %>% 
  mutate(
    offense = case_when(
      offense_code %in% c("12A") ~ "personal robbery",
      offense_code %in% c("13A", "13B") ~ "assault",
      offense_code %in% c("22A") ~ "residential burglary",
      offense_code %in% c("290") ~ "criminal damage",
      TRUE ~ NA_character_
    ),
    year = year(date_single),
    year_mins = difftime(ceiling_date(date_single, unit = "year", 
                                      change_on_boundary = TRUE), 
                         floor_date(date_single, unit = "year"), 
                         units = "mins"),
    crime_mins = difftime(date_single, floor_date(date_single, unit = "year"), 
                          units = "mins"),
    year_angle = (as.numeric(crime_mins) / as.numeric(year_mins)) * 360
  ) %>% 
  filter(!is.na(offense)) %>% 
  select(-year_mins, -crime_mins, -offense_against, -offense_type, 
         -offense_group)
```

We can quickly check the results of this calculation using a histogram of
daily crime counts (this is for all crime and should be interpreted with 
caution):

```{r}
ggplot(crimes) + 
  geom_histogram(aes(year_angle), bins = 365) + 
  facet_wrap(facets = vars(city_name), scales = "free") +
  theme_minimal()
```


# Identify neighbourhoods by clustering

## Convert shapefiles to geopackages

Zillow data are provided as shapefiles. For ease of manipulation, we can first
convert those shapefiles to geopackages. This only needs to be done once or when
the Zillow data has changed.

```{r message=FALSE, warning=TRUE, include=FALSE}
walk(
  dir("../original_data", pattern = "^Zillow.+\\.zip", full.names = TRUE), 
  function (x) {
    
    # unzip shapefiles to temporary directory
    unzip(x, exdir = tempdir())
    
    # set file names
    input_file <- paste0(tempdir(), str_extract(x, "\\/[^\\/]+$")) %>% 
      str_replace("\\.zip$", ".shp")
    output_file <- str_replace(x, "\\.zip$", ".gpkg") %>% 
      str_replace("\\/original_data\\/", "/analysis_data/")
    cat(input_file, "\n", output_file, "\n")
    
    # save shapefile as geopackage
    st_read(input_file) %>%
      st_write(output_file)
    
    # remove temporary files
    dir(tempdir(), pattern = "^Zillow", full.names = TRUE) %>% 
      file.remove()
  }
)
```


## Extract neighbourhoods for CODE cities

```{r}
nbhd <- dir("../analysis_data", pattern = "^Zillow.+\\.gpkg", 
            full.names = TRUE) %>% 
  map_df(st_read, stringsAsFactors = FALSE) %>% 
  rename_all(str_to_lower) %>%
  rename(geometry = geom, neighborhood = name) %>%
  filter(city %in% c("Chicago", "Detroit", "Fort Worth", "Kansas City",
                     "Los Angeles", "Louisville", "New York", "San Francisco",
                     "Tucson", "Virginia Beach")) %>%
  st_set_crs(4326) %>% 
  mutate_at(vars(one_of("state", "county", "city", "neighborhood", "regionid")), 
            as.character)
```

Not all areas are covered by Zillow neighbourhoods. The quickest way to check is
to produce maps of the neighbourhoods in each city.

```{r}
base::split(nbhd, as.factor(nbhd$city)) %>% walk(function (x) {
  this_outline <- tigris::places(cities$fips[cities$name == first(x$city)], 
                                 cb = TRUE, class = "sf") %>% 
    filter(NAME == first(x$city))
  this_plot <- ggplot(x) + 
    geom_sf(data = this_outline, colour = NA, fill = "grey80") +
    geom_sf(aes(fill = neighborhood), show.legend = FALSE) +
    geom_sf(data = this_outline, colour = "black", fill = NA, linetype = 3) +
    labs(title = paste(first(x$city), "neighborhoods")) +
    theme_minimal()
  ggsave(paste0("../output/", first(x$city), " neighborhoods.pdf"), this_plot)
})
```

All of Chicago, Los Angeles, New York and San Francisco, and almost all of 
Lousville and Virginia Beach are covered by Zillow neighbourhoods, so they can
be used for neighbourhood-level analysis.

```{r}
nbhd %>% 
  as_tibble() %>% 
  filter(city %in% c("Chicago", "Los Angeles", "Louisville", "New York",
                     "San Francisco", "Virginia Beach")) %>% 
  group_by(city) %>% 
  summarise(n = n()) %>%
  add_row(city = "Total", n = sum(.$n))
```


## Join Zillow neighbourhoods to crime data

```{r}
crimes <- crimes %>% 
  # convert the crimes object to an SF object, preserving the original lat and 
  # lon columns
  st_as_sf(coords = c("longitude", "latitude"), crs = 4326, remove = FALSE) %>% 
  # join the neighbourhood names (but no other information) to the crimes
  st_join(select(nbhd, neighborhood)) %>% 
  # convert the SF object back to a standard data frame and convert that to a
  # tibble, thereby restoring crimes to its original class
  st_set_geometry(NULL) %>% 
  as_tibble() %>% 
  # remove duplicate rows introduced by st_join() when points fall on the
  # boundary between two neighbourhoods (note summarise_all() does not work on
  # SF objects, so this must be done after the conversion to tibble)
  group_by(uid) %>% 
  summarise_all(first) %>% 
  ungroup() %>% 
  # remove neighbourhood value for cities that are not completely or largely
  # covered by neighbourhoods
  mutate(neighborhood = case_when(
    !city_name %in% c("Chicago", "Los Angeles", "Louisville", "New York",
                 "San Francisco", "Virginia Beach") ~ NA_character_,
    TRUE ~ neighborhood
  ))
```

