---
title: "02 Data preparation"
output: html_notebook
---


# Load CODE crime data

```{r}
crimes <- dir("../original_data", pattern = "_core_", full.names = TRUE) %>% 
  map(read_csv, col_types = cols(
    .default = col_character(),
    uid = col_integer(),
    date_single = col_datetime(format = ""),
    date_start = col_datetime(format = ""),
    date_end = col_datetime(format = ""),
    longitude = col_double(),
    latitude = col_double()
  )) %>% 
  bind_rows() %>% 
  # Virginia Beach data is only available for four years, so is excluded
  filter(city_name != "Virginia Beach") %>% 
  # not all fields are needed, so some a excluded to save memory
  select(city_name, offense_code, offense_type, offense_group, offense_against, 
         date_single, date_start, date_end, longitude, latitude, location_type, 
         location_category, fips_state, fips_county, tract, block_group, block)

crimes %>% 
  mutate(year = year(date_single)) %>% 
  group_by(city_name, year) %>% 
  summarise(n = n()) %>% 
  spread(year, n)
```


# Create circular representations of dates

Circular statistics require dates to be represented as angles showing the 
position of that date within a larger unit. For example, an angle could show
how far through the year, week, day etc a date/time combination falls. Since
we're interested in seasonal variations, we will represent dates as an angle
relative to a year.

```{r}
crimes <- crimes %>% 
  mutate(
    offense = case_when(
      offense_code %in% c("12A") ~ "personal robbery",
      offense_code %in% c("13A", "13B") ~ "assault",
      offense_code %in% c("22A") ~ "residential burglary",
      offense_code %in% c("290") ~ "criminal damage",
      TRUE ~ NA_character_
    ),
    year = year(date_single),
    year_mins = difftime(ceiling_date(date_single, unit = "year", 
                                      change_on_boundary = TRUE), 
                         floor_date(date_single, unit = "year"), 
                         units = "mins"),
    crime_mins = difftime(date_single, floor_date(date_single, unit = "year"), 
                          units = "mins"),
    year_angle = (as.numeric(crime_mins) / as.numeric(year_mins)) * 360
  ) %>% 
  filter(!is.na(offense)) %>% 
  select(-year_mins, -crime_mins, -offense_against, -offense_type, 
         -offense_group)
```

We can quickly check the results of this calculation using a histogram of
daily crime counts (this is for all crime and should be interpreted with 
caution):

```{r}
ggplot(crimes) + 
  geom_histogram(aes(year_angle), bins = 365) + 
  facet_wrap(facets = vars(city_name), scales = "free") +
  theme_minimal()
```


