---
title: "03 Data analysis"
output: html_notebook
---


This file assumes that files `01_packages_and_functions.Rmd` and 
`02_data_prepration.Rmd` have already been run, so that the global environment
contains a tibble of crime data.


# Are seasonal trends each each city consistent across years?

Seasonal analysis often aggregates data from multiple years, and this study 
seeks to do the same. However, this approach is only valid if the seasonal
pattern is the same across years. These tests seek to determine if this 
assumption is valid.

```{r}
crimes %>% 
  filter(!offense %in% c("theft from the person", "theft of motor vehicle")) %>% 
  split(.$offense) %>% 
  # head(1) %>% 
  lapply(function (x) {
    x %>% split(.$city_name) %>% 
      lapply(function (y) {
        z <- split(y$year_angle, y$year) %>% 
          lapply(circular_clock)
        medians <- sapply(z, quantile.circular, probs = c(0.50)) %>% 
          abs_circular()
        # plot(medians, main = paste(first(x$offense), "in", first(y$city_name)), 
        #      ylim = c(0, 360))
        concentrations <- sapply(z, circular_concentration)
        # calculate range of annual median values here
        watson.wheeler.test(z) %>% 
          tidy() %>% 
          mutate(
            type = first(x$offense), 
            city = first(y$city_name),
            median_length = median(sapply(z, length)),
            median_day = angle_to_day(median(medians)),
            median_range = paste(
              round(((max(medians) - min(medians)) / 360) * 365), 
              "days"
            ),
            median_conc = median(concentrations)
          ) %>% 
          select(type, city, median_length, median_day, median_range, 
                 median_conc, statistic, p = p.value)
      }) %>% 
      bind_rows()
  }) %>% 
  bind_rows() %>% 
  mutate(
    median_length = round(median_length),
    median_conc = round(median_conc, digits = 3),
    p_adj = p.adjust(p, "hochberg")
  ) %>% 
  mutate_at(vars(one_of("statistic")), round, 
            digits = 1) %>% 
  mutate_if(str_detect(colnames(.), "^p"), 
            ~ifelse(. < 0.001, "< 0.001", as.character(round(., digits = 3))))
```


# Are seasonal patterns of crime consistent across cities?

We can generate some summary statistics to see if there are differences between
the seasonal distributions of crimes in each CODE city.

```{r}
city_summary <- crimes %>% 
  # filter(city_name == "Fort Worth") %>% 
  split(.$offense) %>% 
  lapply(function (x) {
    cat("\n\nPROCESSING", str_to_upper(first(x$offense)), "\n")
    
    split(x, x$city_name) %>% 
      lapply(function (y) {
        cat("\tfor", first(y$city_name), "\n")
      
        # create a circular object of dates represented as angles
        z <- circular_clock(y$year_angle)
        
        # conduct 10 Rayleigh tests (because the test breaks ties at random)
        # and take the mean values
        rayleigh <- map_df(1:20, ~rayleigh.test(z)[1:2]) %>% 
          summarise_all(mean)
        
        # construct a tibble of statistics
        tibble(
          offenses = nrow(y),
          rayleigh_stat = rayleigh$statistic,
          rayleigh_p = rayleigh$p.value,
          median_angle = median.circular(z) %>% 
            abs_circular(),
          median_day = angle_to_day(median_angle),
          mean_deviation = meandeviation(z),
          concentration = circular_concentration(z),
          seasonal_range = seasonal_range(y$date_single, period = "V")
        )
        
      }) %>% 
      bind_rows(.id = "city")
  }) %>% 
  bind_rows(.id = "offense") %>% 
  mutate(
    rayleigh_p = format_p(p.adjust(rayleigh_p))
  ) %>%
  write_csv("../analysis_data/city_summary.csv", na = "")

city_summary %>% 
  mutate_if(is.integer, format, big.mark = ",") %>% 
  mutate_if(is.numeric, round, digits = 3)
```

The Rayleigh statistics show that in all cases there is significant seasonal
variation in the data, although the concentration and seasonal-range values
suggest this variation is relatively small.

We can plot these results to see that assault and criminal damage are 
concentrated in the summer, robbery in the late summer and burglary in the 
autumn and winter.

```{r}
city_summary %>% 
  ggplot(aes(x = (median_angle / 360) * 365, y = concentration, label = str_sub(city, 0, 3))) +
  # geom_hline(yintercept = 1, colour = "grey70") +
  geom_point() +
  geom_text_repel() +
  facet_wrap(facets = "offense") +
  coord_polar() +
  scale_x_continuous(
    breaks = c(31, 59, 90, 120, 151, 181, 212, 243, 273, 304, 334, 365),
    label = c("J", "F", "M", "A", "M", "J", "J", "A", "S", "O", "N", "D"),
    limits = c(0, 365)
  ) +
  scale_y_continuous(
    limits = c(0, NA)
  ) +
  theme_minimal() +
  theme(
    axis.title.x = element_blank(),
    # axis.line.y = element_blank(),
    # axis.text.y = element_blank(),
    axis.title.y = element_blank(),
    panel.grid.minor.y = element_blank(),
    panel.grid.minor.x = element_blank()
  )
```

We can also see that the differences in concentration between cities are greater 
than the differences between crime types.

```{r}
city_summary %>% 
  ggplot(aes(
    x = offense, 
    y = concentration, 
    group = city, 
    label = ifelse(.$offense %in% c(first(.$offense), last(.$offense)), 
                   paste(str_sub(city, 0, 3), " "), "")
  )) + 
  geom_line() + 
  geom_point() + 
  geom_text(hjust = "right") + 
  lims(y = c(0, NA)) + 
  theme_minimal() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank()
  )
```

In each case, these differences between crime types are significant.

```{r}
crimes %>% 
  split(.$offense) %>% 
  map_df(function (x) {

    y <- split(x$year_angle, x$city_name) %>% 
      map(circular_clock) 
    
    # we run the WW test 20 times because it breaks ties at random
    map_df(1:20, ~tidy(watson.wheeler.test(y))) %>% 
      summarise_if(is.numeric, mean)
      # watson.wheeler.test(z) %>%
      # tidy() %>% 
      

  }, .id = "offense") %>% 
  mutate(p.value = format_p(p.adjust(p.value))) %>%
  rename(p_value = p.value)
```















```{r}
crimes %>% 
  split(.$offense) %>% 
  head(1) %>% 
  lapply(function (x) {
    
    cat("\nProcessing", first(x$offense), "\n")
    y <- split(x$year_angle, x$city_name) %>% 
      lapply(circular_clock)
    cat("\tcreated list of circular objects\n")
    
    medians <- y %>% 
      sapply(function (z) {
        median.circular(ifelse(length(z) > 10000, sample(z, 10000), z))
      }) %>%
      abs_circular()
    cat("\tcalculated medians\n")
    
    concentrations <- sapply(y, function(z) {
      circular_concentration(ifelse(length(z) > 10000, sample(z, 10000), z))
    })
    cat("\tcalculated concentrations\n")
    
    watson.wheeler.test(y) %>%
      tidy() %>%
      mutate(
        type = first(x$offense),
        median_length = median(sapply(y, length)),
        median_day = angle_to_day(median(medians)),
        median_range = paste(
          round(((max(medians) - min(medians)) / 360) * 365),
          "days"
        ),
        median_conc = median(concentrations)
      ) %>%
      select(type, median_length, median_day, median_range, median_conc,
             statistic, p = p.value)

  }) %>%
  bind_rows() %>%
  mutate(
    median_length = round(median_length),
    median_conc = round(median_conc, digits = 3),
    statistic = round(statistic, digits = 1),
    p_adj = p.adjust(p, "hochberg")
  ) %>%
  mutate_if(str_detect(colnames(.), "^p"),
            ~ifelse(. < 0.001, "< 0.001", as.character(round(., digits = 3))))
```

```{r}
crimes %>% 
  filter(offense == "personal robbery") %>% 
  { split(.$year_angle, .$city_name) } %>% 
  lapply(function (x) {
    circular_clock(x) %>% 
      density.circular(bw = bw.nrd.circular(.), adjust = 10) %>% 
      { tibble(x = .$x, y = .$y) }
  }) %>% 
  bind_rows(.id = "city") %>% 
  mutate(x_abs = abs_circular(x)) %>% 
  # sample_n(50) %>% View()
  ggplot(aes(x = x_abs, y = y)) +
  geom_line() +
  coord_polar() +
  facet_wrap(vars(city)) +
  theme_minimal()

```

